Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-08-13 Latex_files/Thesis/chapters/chapter 3/main.tex
updates to ch3

Supersaturated solutions of glycine and water were prepared and stored in
an incubator prior to use. When ready to be studied $15\ \mu L$ was pipetted 
into the solution and $20\ \mu L$ was pipetted onto the cover slip. A Vaterite
sphere was located, trapped, and moved as close to the droplet edge as possible.
After measuring the microsphere's rotational frequency the sphere was left to
rotate for a period of ten minutes after which, if no nucleation event was 
observed the particle was released. The overall results are catalogued in 
table~\ref{tab:Nucleation}
\begin{table}[h!]
	\centering
	\caption{Results from rotating Vaterite within supersaturated solution of $H_2O$ and Glycine. Solubility concentration for Glycine at $16^\circ$ was $C^*=0.2016g/g$}
	\label{tab:Nucleation}
	\begin{tabular}[width=\textwidth]{|c|c|c|c|}
		\hline
		Super Saturation & Particle radius [$\mu$] & $\omega$ [Hz] & Nucleation [$\checkmark/\times$]\\
		\hline
		\multirow{3}*{1.01} & 2.34 & 20.7 & $\times$ \\
		\cline{2-4} & 5.67 & 23.3 & $\times$ \\
		\cline{2-4} & 3.26 & 15.4 & $\times$ \\
		\hline
		\multirow{3}*{1.14} & 1.89 & 1.23 & $\times$ \\
		\cline{2-4} & 3.75 & 3.54 & $\times$ \\
		\cline{2-4} & 4.35 & 4.86 & $\times$ \\
		\hline
		\multirow{3}*{1.4} & 3.47 & 0.00 & $\times$ \\
		\cline{2-4} & 1.59 & 0.00 & $\times$ \\
		\cline{2-4} & 6.24 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.45} & 6.32 & 0.00 & $\times$ \\
		\cline{2-4} & 3.68 & 0.00 & $\times$ \\
		\cline{2-4} & 5.43 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.49} & 4.76 & $0.00$ & $\times$ \\
		\cline{2-4} & 7.27 & $0.00$ & $\times$ \\
		\cline{2-4} & 1.52 & $0.00$ & $\times$ \\
		\hline
	\end{tabular}
\end{table}

@@

\begin{align}
	u_r(r)=-|v(t)|^2cos(\theta)\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
\end{align}

Where $\theta$ is the angle from the direction of movement to the point
you wish to measure, and $r$ is the radial distance to that point. Again
taking the partial derivative we can get the shear rate for a particle moving 
through the fluid:

\begin{align}
	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = |v(t)|^2cos(\theta)\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
\end{align}

Moving a silica bead along an circular path can generate significant fluid 
flow around a larger volume compared to comparable microspheres.

@@

\subsection{Stationary beam}
\label{sec:stationary}
Consider below in Fig.~\ref{fig:stationary_beam} the frames taken from a nucleation 
event in supersaturated glycine solution ($S=1.03$), the beam is a stationary being 
$\approx3.5 \mu m$ from the droplet edge. After a period of roughly 5 minutes a 
nucleus forms at the trap focus, growing quickly (growth rate was approximated using
imageJ to be on the order of $700\ \mu m^2/min$) from the focal point of the trap 
until after roughly $6$ seconds the crystal escapes. Comparing to previous literature
using optical tweezers shows that the growth rate is only loosely connected to the 
solutions supersaturation. Local conditions play a much larger role in the growth 
rate than just the concentration \cite{Flannigan2023}. A likely reason that the trap 
is escaped is due to the fact that crystal is far too large to be held in place and 
is in fact still growing as the solution is supersaturated. The key take away to 
remember is that the beam has no real influence over the crystal shape, instead 
it grows outward from the trap focus.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.7\linewidth]{frames_no_beam_movement.pdf}
	\caption{Laser induced nucleation at the edge of a droplet of supersaturated 
	glycine solution. (b) shows the first instance of a crystal nucleus, growing 
	quickly through (c)-(e) until after $6.665\ s$ the crystal begins to escape the trap.}
	\label{fig:stationary_beam}
\end{figure}

@@

Interestingly the galvano-mirror allows the trap to impart a slight torque 
on the crystal, as shown in fig.~\ref{fig:eliptical_beam_1}(c) and (d), 
where even though the crystal is not directly in the trap focus it rotates 
in the $x-y$ plane and gets trapped again at a corner. The rotation could 
not be due to fluid flow close to the surface of the crystal as trap should 
have no influence on fluid molecules smaller than a few hundred nano-meters. 
In figs.\ref{fig:eliptical_beam_1}(e) and (f), the crystal growth becomes 
localised to the corner. The area growth rate between figures \ref{fig:eliptical_beam_1}(a) and (d) was approximated using imageJ at $45.03\ 
\mu m^2 /min$, where as between figures \ref{fig:eliptical_beam_1}(e) and 
(f) the growth rate at that particular edge was estimated at $42.10\ \mu m^2/min$. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.9\linewidth]{frames_eliptical_beam.pdf}
	\caption{Laser induced nucleation at the edge of a droplet of 
		supersaturated glycine solution. (b) shows the first instance 
		of a crystal nucleus, growing quickly through (c)-(e) until 
		after $6.665\ s$ the crystal begins to escape the trap.}
	\label{fig:eliptical_beam_1}
\end{figure}

@@

\subsection{Direct trapping of Glycine clusters}
\label{sec:clusters}
One common aspect involving these droplets is the fact when brought to the laser 
focus the droplets would nucleate immediately \cite{Liao2022}. We made up a 
solution similar to Sec~\ref{sec:stationary} but without any silica droplets, 
we once again focused the beam close to the droplet edge. Again after a few 
minutes of irradiation we noted the presence of droplets, because no silica had 
been added these droplets had to be from the glycine solution. Trapping these

It has been shown that polarisation can influence the polymorphic control for 
simple salts \cite{Garetz1996, Garetz2002} but no research has been done into 
the influence of a moving beam front on the growth of a newly formed nucleus.

\section{Influence of a moving beam front on seed crystals}
To see if the above phenomena had any direct influence on a

>>>>

Supersaturated solutions of glycine and water were prepared and stored in
an incubator prior to use. When ready to be studied $15\ \mu L$ was pipetted 
into the solution and $20\ \mu L$ was pipetted onto the cover slip. A Vaterite
sphere was located, trapped, and moved as close to the droplet edge as possible.
After measuring the microsphere's rotational frequency the sphere was left to
rotate for a period of ten minutes after which, if no nucleation event was 
observed the particle was released. The overall results are catalogued in 
table~\ref{tab:Nucleation}
\begin{table}[h!]
	\centering
	\caption{Results from rotating Vaterite within supersaturated solution of $H_2O$ and Glycine. Solubility concentration for Glycine at $16^\circ$ was $C^*=0.2016g/g$}
	\label{tab:Nucleation}
	\begin{tabular}[width=\textwidth]{|c|c|c|c|}
		\hline
		Super Saturation & Particle radius [$\mu m$] & $\omega$ [Hz] & Nucleation [$\checkmark/\times$]\\
		\hline
		\multirow{3}*{1.01} & 2.34 & 10.4 & $\times$ \\
		\cline{2-4} & 5.67 & 9.63 & $\times$ \\
		\cline{2-4} & 3.26 & 8.46 & $\times$ \\
		\hline
		\multirow{3}*{1.14} & 1.89 & 1.23 & $\times$ \\
		\cline{2-4} & 3.75 & 3.54 & $\times$ \\
		\cline{2-4} & 4.35 & 4.86 & $\times$ \\
		\hline
		\multirow{3}*{1.4} & 3.47 & 0.00 & $\times$ \\
		\cline{2-4} & 1.59 & 0.00 & $\times$ \\
		\cline{2-4} & 6.24 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.45} & 6.32 & 0.00 & $\times$ \\
		\cline{2-4} & 3.68 & 0.00 & $\times$ \\
		\cline{2-4} & 5.43 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.49} & 4.76 & $0.00$ & $\times$ \\
		\cline{2-4} & 7.27 & $0.00$ & $\times$ \\
		\cline{2-4} & 1.52 & $0.00$ & $\times$ \\
		\hline
	\end{tabular}
\end{table}

@@

\begin{align}
	u_r(r)=-|v(t)|^2\left(1-\frac{3R}{2r}+\frac{R^3}{2r^3}\right)
\end{align}

Where $r$ is the radial distance to that point. Again taking the partial 
derivative we can get the shear rate for a particle moving through the fluid:
\begin{align}
	\dot{\gamma}(r) = \left| \frac{\delta u_r(r)}{\delta r}\right| = |v(t)|^2\left(\frac{3R}{r^2} -\frac{2R^3}{r^4} \right)
	\label{eq:galvo_shear}
\end{align}

Moving a silica bead along an circular path can generate significant fluid 
flow around a larger volume compared to comparable micro-rotors. Using \eqref{eq:galvo_shear} we estimated the shear rate that the surrounding fluid 
would experience at varying speeds. The maximum speed of $7000\ \mu m s^{-1}$
equates to moving the silica bead around a circular path with a frequency of 
$100\ Hz$.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{galvano_shear_rate.png}
	\caption{Shear rate generated by a silica microsphere ($a = 1.57\ \mu m$) 
		in bulk fluid moving at different speeds. The shear rate is calculated 
		using \eqref{eq:galvo_shear}, with the assumption that the bead is moving
		in a circular path and so the speed is constant through out its path.}
	\label{fig:galvano_shear}
\end{figure}

From figure~\ref{fig:galvano_shear} it is clear that not only is a galvano 
mirror a better option for generating high shear rates but also over a larger 
volume.

@@

\subsection{Stationary beam}
\label{sec:stationary}
Consider below in Fig.~\ref{fig:stationary_beam} the frames taken from a nucleation 
event in supersaturated glycine solution ($S=1.03$), the beam is a stationary being 
$\approx3.5 \mu m$ from the droplet edge. After a period of roughly 5 minutes a 
nucleus forms at the trap focus, growing quickly (growth rate was approximated using
imageJ to be on the order of $700\ \mu m^2/min$) from the focal point of the trap 
until after roughly $6$ seconds the crystal escapes. Comparing to previous literature
using optical tweezers shows that the growth rate is only loosely connected to the 
solutions supersaturation. Local conditions play a much larger role in the growth 
rate than just the concentration \cite{Flannigan2023}. A likely reason that the trap 
is escaped is due to the fact that crystal is far too large to be held in place and 
is in fact still growing as the solution is supersaturated. The key take away to 
remember is that the beam has no real influence over the crystal shape, instead 
it grows outward from the trap focus.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.63\linewidth]{frames_no_beam_movement.pdf}
	\caption{Laser induced nucleation at the edge of a droplet of supersaturated 
	glycine solution. (b) shows the first instance of a crystal nucleus, growing 
	quickly through (c)-(e) until after $6.665\ s$ the crystal begins to escape the trap.}
	\label{fig:stationary_beam}
\end{figure}

@@

Interestingly the galvano-mirror allows the trap to impart a slight torque 
on the crystal, as shown in fig.~\ref{fig:eliptical_beam_1}(c) and (d), 
where even though the crystal is not directly in the trap focus it rotates 
in the $x-y$ plane and gets trapped again at a corner. The rotation could 
not be due to fluid flow close to the surface of the crystal as trap should 
have no influence on fluid molecules smaller than a few hundred nano-meters. 
In figs.\ref{fig:eliptical_beam_1}(e) and (f), the crystal growth becomes 
localised to the corner. The area growth rate between figures \ref{fig:eliptical_beam_1}(a) and (d) was approximated using imageJ at $45.03\ 
\mu m^2 /min$, where as between figures \ref{fig:eliptical_beam_1}(e) and 
(f) the growth rate at that particular edge was estimated at $42.10\ \mu m^2/min$. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.85\linewidth]{frames_eliptical_beam.pdf}
	\caption{Frames from a longer video depicting the growth of a nucleus using 
		a moving beam. Initially the crystal shape is amorphous (a) but eventually
		reaches a more regular shape (b). This crystal is still influenced by the
		optical trap as even when not directly irradiated by the laser the crystal
		rotates between (c) and (d). When the laser is focused on a corner the 
		crystal growth is localised to that region, resulting in an elongated 
		section forming between frames (e) and (f).}
	\label{fig:eliptical_beam_1}
\end{figure}

@@

\subsection{Direct trapping of Glycine clusters}
\label{sec:clusters}
One common aspect involving these droplets is the fact when brought to the laser 
focus the droplets would nucleate immediately \cite{Liao2022}. A solution similar 
to Sec~\ref{sec:stationary} was made up, but without any silica droplets. Once 
again the beam was focused close to the droplet edge, this time the galvano mirror
was scanning a circular path (as shown in figure~\ref{fig:cluster_trapping}). 
After a few minutes of irradiation droplets were seen entering the camera frame, 
because no silica had been added these droplets had to be from the glycine solution. Trapping individual droplets did not result in immediate nucleation even after 
several minutes being trapped. Trying to bring two droplets together resulted in nucleation between the two droplets, compared to \ref{sec:stationary} \& 
\ref{sec:moving} the growth is much slower, taking nearly 40 seconds before the 
crystal structure becomes clearer.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.9\linewidth]{cluster_trapping.pdf}
	\caption{Frames from a longer video demonstrating the trapping of a glycine 
		droplet. Solution is undersaturated glycine and water (S = 0.86), with the 
		laser power is set at $750\ mW$. (a) shows a trapped droplet (outlined in 
		black) being brought into contact with a larger droplet. (b) upon contact 
		a nucleus can be seen between the two droplets. The growth is rather slow 
		with the crystal having no clear defined morphology through (c) and (d). 
		Between frames (e) and (f) the larger droplet finally joins the main crystal.}
	\label{fig:cluster_trapping}
\end{figure}

The fact that these droplets can be trapped indicates they must have a higher 
refractive index than the surrounding solution. However, getting an exact 
measurement of the

\section{Influence of a moving beam front on seed crystals}
\textcolor{red}{Need to write up these results. To summarise the crystal 
	dissolved around the laser focus instead of growing as the seed crystals 
	are in the solution and not close to the air-solution interface}
